### 2024-04-09 paper/01_motivation.tex

Machine learning techniques have emerged as a powerful tool in the toolkit of
materials scientists. While they are often used to make predictions on
the properties of materials or find new materials with certain properties, an increasingly interesting domain is the automated analysis of raw experimental measurements
guided by machine learning\supercite{radovicMachineLearningEnergy2018}.

With the advent of high-throughput experiments, the amount of gathered data is
vast and the analysis often becomes a bottleneck in the processing pipeline
\supercite{rahmanianEnablingModularAutonomous2022}. Powder X-ray diffraction
(XRD) is an important measurement technique used to obtain structural
information from polycrystalline samples\supercite{harrisContemporaryAdvancesUse2001}. 
The diffractograms are an
information-dense fingerprint of the structure of the material. However,
analyzing these diffractograms is not an easy task\supercite{holderTutorialPowderXray2019}. 
Full structure solutions and
Rietveld refinement take time and require expert knowledge, both about the
analysis technique and the materials class at hand. This is not feasible in high-throughput experiments on a larger
scale. Therefore, the question arises whether it is possible to automatically 
analyze powder diffractograms with machine learning models trained on large
amounts of data, making it possible to run inference almost instantaneously.

During the last few years, there have been several studies tackling this
objective by applying machine learning models to various tasks concerning the
analysis of powder diffractograms such as phase
classification\supercite{leeDeeplearningTechniquePhase2020,maffettoneCrystallographyCompanionAgent2021,schuetzkeEnhancingDeeplearningTraining2021,szymanskiProbabilisticDeepLearning2021,wangRapidIdentificationXray2020},
phase fraction determination\supercite{leeDatadrivenXRDAnalysis2021}, space
group
classification\supercite{parkClassificationCrystalStructure2017,oviedoFastInterpretableClassification2019,zalogaCrystalSymmetryClassification2020,vecseiNeuralNetworkBased2019,suzukiSymmetryPredictionKnowledge2020,chakrabortyDeepCrystalStructure2022},
machine-learning-guided Rietveld
refinement\supercite{ozakiAutomatedCrystalStructure2020,fengMethodArtificialIntelligence2019},
extraction of lattice
parameters\supercite{dongDeepConvolutionalNeural2021,chitturiAutomatedPredictionLattice2021,chakrabortyDeepCrystalStructure2022,
habershonPowderDiffractionIndexing2004} and crystallite
sizes\supercite{dongDeepConvolutionalNeural2021,chakrabortyDeepCrystalStructure2022},
and also novelty detection based on unsupervised techniques\supercite{bankoDeepLearningVisualization2021}.
Since an abundant source of experimental diffractograms is hard to come by, most
applications train their models on simulated diffractograms from the Inorganic
Crystal Structure Database (ICSD)\supercite{bergerhoff1987}, which contains a total of 272\,260 structures (October 2022).

\citeauthor{leeDeeplearningTechniquePhase2020} used a deep convolutional neural
network (CNN) trained on a large dataset of multiphase compositions from the
quaternary \ce{Sr-Li-Al-O} pool to classify present phases in the
diffractogram\supercite{leeDeeplearningTechniquePhase2020}. In a follow-up
study, they further showed good results for phase fraction inference in the
quaternary \ce{Li-La-Zr-O} pool\supercite{leeDatadrivenXRDAnalysis2021}.
\citeauthor{schuetzkeEnhancingDeeplearningTraining2021} performed phase
classification on iron ores and cement compounds and used data augmentation with
respect to lattice parameters, crystallite sizes, and preferred orientation
\supercite{schuetzkeEnhancingDeeplearningTraining2021}. They showed that
especially the lattice parameter variations enhance the classification accuracy
significantly.

Instead of the analysis of phase composition,
\citeauthor{dongDeepConvolutionalNeural2021} performed regression of scale
factors, lattice parameters, and crystallite sizes in a five-phase catalytic
materials system\supercite{dongDeepConvolutionalNeural2021}. In contrast to
supervised tasks, \citeauthor{bankoDeepLearningVisualization2021} used a
variational autoencoder to visualize variations in space group, preferred
orientation, crystallite size, and peak
shifts\supercite{bankoDeepLearningVisualization2021}.
\citeauthor{parkClassificationCrystalStructure2017} used a deep CNN to classify
space groups of single-phase diffractograms, reaching a test accuracy of 81.14\%
on simulated diffractograms.\supercite{parkClassificationCrystalStructure2017} 
However, as we will show later in this paper, this accuracy is highly overestimated and drops to 56.1\% when test splits are designed in a way to reduce data leakage in non-IID datasets such as the ICSD.
\citeauthor{vecseiNeuralNetworkBased2019}\supercite{vecseiNeuralNetworkBased2019} developed
a similar approach and applied their classifier to experimental diffractograms
from the RRUFF mineral database\supercite{lafuentePowerDatabasesRRUFF2015},
reaching an experimental test accuracy of 54\%.

While the ICSD contains a large number of structures spanning many different
classes of materials, it still falls short in size, distribution, and generality
compared to the datasets used to train very large state-of-the-art models of
other fields such as computer vision.
%for image classification (ImageNet
%challenge\supercite{russakovskyImageNetLargeScale2015}). 
Furthermore, the ICSD
database is highly imbalanced with respect to space groups, as can be seen
in the histogram in Figure~\ref{fig:histogram_dist_spgs}. This makes the classification of
space groups more difficult, as shown and discussed by
\citeauthor{zalogaCrystalSymmetryClassification2020}
\supercite{zalogaCrystalSymmetryClassification2020}. The ICSD also contains a
limited number of different structure types that may not adequately represent
the crystal structures analyzed in future experiments.

To overcome these shortcomings, we propose to train machine learning models on
diffractograms simulated from synthetic crystal structures randomly generated
based on the symmetry operations of the space groups. This makes it possible to
train on structures with new structure types not present in the ICSD. We used
the crystals from the ICSD only to determine vague statistics guiding the random
generation and for calculating the test accuracy. Our approach goes one step
further than classical data augmentation by fully detaching itself from the
individual entries in the ICSD database.
\replaced{The generated synthetic crystals form a training dataset that includes stable ICSD crystal structures, unstable crystal structures,
but also stable structures that are not yet present in the ICSD. By training a model on the full dataset, we can also expect improvements on the unknown stable crystal structures.}{It is our hypothesis that when trying
to solve more general tasks concerning the extraction of structural details from
powder diffractograms, the crystals from which the training dataset is simulated
do not need to be stable or chemically viable.} \replaced{Furthermore, we}{We} propose viewing the problem
\deleted{simply} as a mathematical task of getting back some of the real-space information
leading to given powder X-ray diffractograms. \added{Therefore, even the unstable structures included in our generated dataset will help to learn to classify the stable structures.}

Here, we applied this approach to the classification of the crystal symmetry,
namely the space group. The space group is usually one of the first structural
pieces of information needed after synthesizing a new material. This task is
well-suited to showcase the strengths of using a synthetic dataset and to
benchmark it. We further show the results of using our methodology to infer space group labels of an experimental dataset.

We embedded our synthetic generation algorithm in a framework with
distributed computing capabilities to generate and simulate diffractograms on
multiple nodes in parallel using the \emph{Python} library \emph{Ray}
\supercite{moritzRayDistributedFramework2018a}. In contrast to the traditional
approach of generating a simulated dataset before training, we used this
distributed computing architecture to build an infinite stream of \replaced{synthetically}{randomly}
generated and simulated diffractograms to perform batch-wise online learning.
This increases the generalization performance, eliminates the problem of
overfitting, and allows very large models to be trained.