### 2024-04-09 paper/02_existing_datasets.tex

\subsection{Generating synthetic crystals} \label{sec:generate_crystals}

    To generate synthetic crystals, we randomly place atoms on the Wyckoff
    positions of a given space group \added{following the Wyckoff occupation probabilities extracted from the ICSD} and
    then apply the respective symmetry operations. The algorithm is
    explained in the following (see also Figure~\ref{fig:generation_and_distributed}a for a flow diagram of the algorithm).
    We only explain the most important steps, details can be found in Section~S1 of the SI.

    \begin{enumerate}
        \itemsep0em
        \item \replaced{Sampling of a space group from the space group distribution of the ICSD.}{Random selection of a space group. We follow the space group distribution of the ICSD to allow comparison with previous work.}
        \item Sample unique elements of the crystal and their number of repetitions in the asymmetric unit.
        \item Place atoms onto the Wyckoff positions and draw uniform coordinates for each.
        \item Draw lattice parameters from a \replaced{kernel density estimate}{KDE} based on the ICSD.
        \item Apply space group symmetry operations.
    \end{enumerate}

    Parts of this algorithm were inspired by the generation algorithm of the
    \emph{Python} library \emph{PyXtal}
    \supercite{fredericksPyXtalPythonLibrary2021}. 
    We only keep generated crystals for training if the
    conventional unit cell volume is below \SI{7000}{\angstrom\cubed} and if
    there are less than 100 atoms in the asymmetric unit. 
    We did not employ any form of
    distance checks on the coordinates, as we found this to have no meaningful
    impact on space group classification accuracy. We only prevented the
    algorithm from placing more than one atom onto a Wyckoff position that does not
    have a degree of freedom. We also did not use partial occupancies. We chose
    this algorithm for its simplicity and its capability to reproduce many
    important characteristics of ICSD crystals adequately (see Section~
    \ref{sec:synth_distribution}).

    \begin{figure}[h]
    \centering
    \includegraphics{./figures/distribution_spgs_arrows.pdf}
    \caption{Distribution (logarithmic scale in blue, linear scale in red)
    of space groups in the ICSD. Space groups are sorted by count\added{ (see Figure S13 in the SI for the distribution without sorting by count)}. The population of the space groups varies by
    multiple orders of magnitude, showing that the ICSD is a highly imbalanced
    dataset regarding space groups. The space groups excluded due to insufficient
    statistics are visualized with black stripes. \replaced{The histogram displays the distribution of the full ICSD, while the exclusion of space groups that do not contain enough samples is based on the statistics dataset (which does not include the test dataset, see Section~\ref{sec:dataset_split}) that we used to guide the random crystal generation. Therefore, the excluded space groups are not exactly the first 85 counted from the left.}{Since the exclusion is based on 
    the statistics dataset and this histogram shows the full ICSD,
    the excluded space groups are not exactly the first 85 counted from the left.}}
    \label{fig:histogram_dist_spgs}
    \end{figure}

    For some space groups, there are not enough crystals in the ICSD to form a
    representative \replaced{kernel density estimate}{KDE} for the volume or to calculate suitable occupation
    probabilities for individual Wyckoff positions. Therefore, we chose to only
    perform the classification on space groups with 50 or more crystals
    available in the statistics dataset we used to extract the probabilities
    (see Section~\ref{sec:dataset_split}), resulting in the exclusion of 85
    space groups (see Figure~\ref{fig:histogram_dist_spgs}). A classifier trained directly on ICSD data of all space groups will likely not be able to properly identify these space
    groups containing very few samples.

    If a similar performance for all space groups is desired, a uniform
    distribution of space groups in the training dataset is needed. This is
    trivially possible with our synthetic approach, in contrast to training
    directly on the ICSD, where weighting, over-, or undersampling methods are
    needed \supercite{sunClassificationImbalancedData2009}. To allow a direct
    and fair comparison between our approach and the original approach of
    training directly on ICSD entries, we still followed the same distribution
    of space groups of the ICSD in our synthetic training dataset. This
    eliminates the problem that the effective number of total space groups is
    smaller when training on a highly imbalanced dataset, making it easier to
    reach high accuracies.

    \added{Our choice of not sampling the space groups uniformly and using
    general statistics extracted from the ICSD to guide the crystal generation
    algorithm further builds upon the hypothesis that future crystals will
    roughly follow the more general statistics already present in the ICSD. With
    the chosen crystal generation algorithm we tried to find a middle ground
    between being much more general than using the ICSD crystals directly and
    not being too general such that it is very hard to extract structural
    information at all.}

    \begin{figure}[!htb] 
    \centering
    \includegraphics{./figures/combined_generation_and_distributed_system.pdf}
    \caption{a) Flowchart of how the generation algorithm produces synthetic
    crystals. Atoms are independently placed on the Wyckoff positions and random
    coordinates are drawn. b) Overview of the distributed computing system
    implemented using the \emph{Python} library
    \emph{Ray}\supercite{moritzRayDistributedFramework2018a}. Two compute nodes
    (that generate and simulate diffractograms) are connected to the \emph{Ray}
    head node using a \emph{Ray queue} object.}
    \label{fig:generation_and_distributed}
    \end{figure}

    \subsection{Simulating diffractograms} \label{sec:simulation}

    To simulate powder X-ray diffractograms based on the generated crystals, we used
    the implementation found in the \emph{Python} library \emph{Pymatgen}
    \supercite{ongPythonMaterialsGenomics2013}. We optimized the simulation code
    using the LLVM just-in-time compiler \emph{Numba}
    \supercite{lamNumbaLLVMbasedPython2015}. This increases the performance of
    the main loop over the reciprocal lattice vectors of the crystal
    significantly and makes the continuous simulation while training (discussed
    in the next section) possible.

    We used the wavelength \SI{1.5406}{\angstrom} (\replaced{Cu K$\alpha_1$}{Cu K$_\alpha$} line) to
    simulate all diffractograms. The obtained peaks were further broadened with
    a Gaussian peak profile to form the full diffractogram. To obtain the peak
    widths, we used the Scherrer
    equation\supercite{gilmoreInternationalTablesCrystallography2019}
    \begin{equation}
        \beta=\frac{K \lambda}{L \cos \theta} \text{,} \label{eq:scherrer}
    \end{equation}
    where $ \beta $ is the line broadening at half maximum intensity (on the
    $2\theta$-scale), $K$ is a shape factor, $\lambda$ is the wavelength, and
    $L$ is the (average) thickness of crystallites. We drew crystallite sizes
    from the range $ \left[ 20, 100 \right] $ \si{\nano \metre} and used 
    $K=0.9$.

    Diffractograms were generated in the range $ 2 \theta \in \left[ 5, 90
    \right] \si{\degree} $ with step size \SI{0.01}{\degree}. After generating
    each diffractogram, it was rescaled to fit in the intensity range $ \left[
    0,1 \right]$. \added{In Figure S9 of the SI we show an exemplary diffractogram simulated 
    from the ICSD, Figure S10 shows an exemplary diffractogram simulated from a synthetic 
    crystal.}

    \subsection{Continuous generation of training data}\label{sec:online_learning}

    Typically, machine learning models are trained with a fixed dataset predefined at the beginning of training. Sometimes, data augmentation
    is applied to further increase the effective size of this dataset. In
    contrast to that, we generated our dataset on-the-fly, parallel to model training. The
    main advantage of using this approach compared to a fixed-size dataset is
    the eliminated possibility to overfit to individual diffractograms since
    every diffractogram is only used once. Furthermore, not having to
    pre-simulate a dataset before training makes this approach more flexible
    when changing simulation parameters.

    We used a distributed architecture on multiple nodes using the \emph{Python}
    framework \emph{Ray} \supercite{moritzRayDistributedFramework2018a}, which enabled the training on 1-2 GPUs and simultaneous generation of training data on more than 200 CPU cores (see Figure~\ref{fig:generation_and_distributed}b and SI Section~S2.2). 
    Depending on
    the model size and corresponding training speed, this setup allows
    training with up to millions of unique diffractograms per hour.

    \subsection{Dataset split} \label{sec:dataset_split}

    The ICSD database contains many structures that are very similar with
    slightly different lattice parameters and coordinates. For example, there
    are 25 entries for \ce{NaCl} (October 2022). Furthermore,
    there are 3898 entries that have the same structure type as \ce{NaCl} and
    thus also similar powder diffractograms. If some of them appear in the
    training dataset and some in the test dataset, the classification will be
    simplified to recognizing the structure type or structure. In that case, the
    test set accuracy will not represent the true generalization performance of the neural
    network. To quantify the true generalization performance, we split the
    dataset in such a way that the same structure type appears either only in
    the training or in the test dataset. We used the structure type definitions
    provided by the ICSD. The obtained accuracy on the test dataset reflects the
    accuracy of our network when being used on a novel sample with a structure
    type not yet present in the ICSD database.

    We want to emphasize that the used test split is very important for the task of
    space group classification and not a trivial choice. The ICSD contains many
    subtypes of structure types (for example, subtypes of perovskites), which we
    regarded as separate structure types in our split. Considering the subtypes
    as the same structure type may also be a viable option when performing the
    split. A combination of a split based on structure type and sum formula or
    similar approaches are also possible. \deleted{However, we think that our choice of
    splitting the dataset is able to effectively measure the generalization
    performance while still being relatively simple.}

    \added{Depending on the experimental setting, it further might make more sense in some cases to not do a structure type-based split. If the likelihood of finding structures similar to already-discovered structure types in the planned experiment is high, training should definitely include those structure types to evaluate the performance of the model. However, in a pure discovery setting, new structure types can appear. To evaluate the expected model performance in this scenario and thus quantify the true generalization error to unseen data, we chose the most strict structure type-based split.}

    We divided the ICSD (database version 2021, June 15) in a 70:30 split. For
    our synthetic crystal approach, the 70\% part (\replaced{which we
    call}{called} statistics or training dataset) was only used to create the
    \replaced{kernel density estimates}{KDEs} and to calculate the Wyckoff
    occupation probabilities needed for the generation algorithm. Since we can
    judge the performance of the synthetic generation algorithm by comparing the
    training accuracy (on synthetic crystals) with the accuracy tested on
    diffractograms simulated directly from the statistics dataset, an additional
    validation dataset was not needed. For comparison with the original approach
    of directly training on ICSD
    crystals\supercite{parkClassificationCrystalStructure2017}, we simulated
    crystals directly from the statistics dataset and trained on them.

    Analogous to the synthetic generation, we only used crystals with a
    conventional unit cell volume below \SI{7000}{\angstrom\cubed} and with less
    than 100 atoms in the asymmetric unit for the statistics and test dataset.
    This covers $ \approx 94\% $ of the ICSD crystals.

    \subsection{Models and \added{computational} experiments} \label{sec:models_and_experiments}

    \subsubsection*{Models}
    We will briefly introduce the models we used for the classification
    of space groups. A more detailed description can be found in the SI
    Section~S2.1.

    As a baseline, we first used the CNN models introduced by
    \citeauthor{parkClassificationCrystalStructure2017}\supercite{parkClassificationCrystalStructure2017}.
    They used three models, one for the classification of crystal systems
    (``parkCNN small''), one for extinction groups (``parkCNN medium''), and one
    for space groups (``parkCNN big''). All models have three convolution layers
    with two hidden fully connected layers and one output layer. The
    three models differ in the number of neurons in the hidden fully connected
    layers, increasing the number of model parameters with the number of target
    labels. Here, we only used the models ``parkCNN medium'' and ``parkCNN big''
    and applied both to the classification of space groups. When using ICSD
    crystals to train the ``parkCNN'' models, dropout was used, while the
    training of the ``parkCNN'' models on synthetic crystals did not use
    dropout.

    Since the approach of using an infinite stream of generated training data
    eliminates the problem of overfitting, we further used deeper models with a
    higher number of model parameters. For this, we used the deep convolutional
    neural networks ResNet-10, ResNet-50, and ResNet-101, which were introduced
    by
    \citeauthor{heDeepResidualLearning2016}\supercite{heDeepResidualLearning2016}
    in 2015.

    Details of the machine learning setup can be found in the SI Section~2.2. Overall, our setup allowed us the training of models over up to 2000 epochs with more than $100\,000$ unique, newly generated crystals and corresponding diffractograms in each epoch \added{(see the upper x-axis of Figure \ref{fig:training_curve})}.

    \subsubsection*{\replaced{Computational experiments}{Experiments}}

    We performed two sets of experiments to evaluate our new dataset split as well as our synthetic crystal generation approach and compare it to state-of-the-art models in literature: Firstly, we trained and tested models on ICSD crystals only, and secondly, we trained on synthetic crystals and tested on ICSD crystals.

    In particular, we first performed an experiment with the ``parkCNN medium''
    model trained directly on the diffractograms simulated from the ICSD
    statistics dataset with a fully random train-test split (similar to
    \supercite{parkClassificationCrystalStructure2017}), instead of splitting by
    the structure type of the crystals. This experiment makes a comparison of
    the two different methods of train-test split possible. We then trained the
    ``parkCNN big'' model using the structure type-based split, again directly
    on ICSD diffractograms. We further repeated the same experiment using the
    smaller model ``parkCNN medium'' to resolve potential overfitting to the
    ICSD diffractograms.

    For the experiments performed on our continuously generated
    dataset based on synthetic crystals, we used the structure type-
    based split. As discussed in Section~\ref{sec:dataset_split}, the training / 
    statistics dataset was only used to extract more general statistics, such as the 
    element distribution.
    First, we trained the ``parkCNN big'' model. For each
    batch, we generated $ 435 $ random crystals and simulated two diffractograms
    with different crystallite sizes for each of them. This resulted in the
    batch size of $ 870 $. 
    Since our synthetic crystal generation algorithm yields an infinite stream
    of unique diffractograms to train on, using much larger models than for the
    fixed ICSD dataset is possible without overfitting. We performed experiments
    for the ResNet-10, ResNet-50, and ResNet-101 models. Instead of generating
    two diffractograms with different uniformly sampled crystallite sizes for each generated crystal (as
    we did for the ``parkCNN big'' model), we now created only one diffractogram
    for each of the $ 145 $ crystals generated for one batch. This is due to 
    the slower training of the ResNet models, which means that reusing the same 
    diffractogram with different crystallite sizes is not necessary to generate
    training data fast enough.

    To obtain the
    highest-possible ICSD test accuracy, we further applied the square root
    function as a preprocessing step to the input diffractograms of the
    network when using the ResNet models. This was suggested by
    \citeauthor{zalogaCrystalSymmetryClassification2020}
    \supercite{zalogaCrystalSymmetryClassification2020} and in their case
    improved classification accuracy by approximately 2 percentage points. Some
    initial tests suggested that this approach also yields a higher accuracy in our
    case, so we used this preprocessing step to train the ResNet
    models.

    While we focused mainly on the methodology of using synthetic crystals to
    extract structural information from powder diffractograms, we also show some
    initial steps toward applying our methods to experimental data. We used the
    publicly available RRUFF mineral database
    \supercite{lafuentePowerDatabasesRRUFF2015} which provides experimental
    measurements, including powder diffractograms \added{(see Figure S11 in the
    SI for an exemplary diffractogram from the RRUFF)}. In order to imitate
    experimental diffractograms, we added Gaussian additive and multiplicative
    noise \added{(similar to
    \supercite{vecseiNeuralNetworkBased2019,szymanskiProbabilisticDeepLearning2021})}
    and a background function based on samples from a Gaussian process to our
    simulated diffractograms. Furthermore, we added a small amount of an
    impurity phase to each diffractogram. Details about the experimental data
    generation protocol can be found in the SI Section~S4\added{, Figure S12
    shows an exemplary synthetic diffractogram with noise, background and an
    impurity phase.} Using the ResNet-50 model, we performed two experiments for
    experimental data, one with the mentioned impurity phase, and one without.