### 2024-04-09 paper/_main.tex

\documentclass[a4paper]{article}

\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[super,sort\&compress,comma]{natbib}
\renewcommand{\supercite}[1]{\cite{#1}}

%% Language and font encodings
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{csquotes}

%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

%% Useful packages
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue}
\usepackage{authblk}
\usepackage{multirow}
\usepackage{nomencl}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{textcomp} 
\usepackage{amsfonts}
\usepackage{outlines} %%same
\usepackage{subfig}
\usepackage[version=4]{mhchem}
\usepackage{siunitx}
\usepackage{bm}
\usepackage[normalem]{ulem} %%temporary, for organization purposes

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage[final]{changes}

\interfootnotelinepenalty=10000
\title{Towards a free high quality powder X-ray diffraction dataset}

\author[1,2]{Daniel Hollarek}
\author[1,2]{Henrik Schopmans}
\author[1,2]{Pasc1al Friederich}
\author[2]{Ben Breitung}
\author[2]{Simon Schweidler}
\author[3]{Andrea Hodge}
\author[3]{Adie Alwen}
\author[4,5]{Carolin Sutter-Fella}
\author[4]{Mriganka Singh}
\author[4]{Tim Kodalle}
\author[4]{Maged Abdelsamie}

% I made the affiliations smaller because there's a lot of them
\renewcommand\Affilfont{\small}

\affil[1]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology,\protect \\
Engler-Bunte-Ring 8, 76131 Karlsruhe, Germany}
\affil[2]{Institute of Nanotechnology, Karlsruhe Institute of Technology,\protect \\ Hermann-von-Helmholtz-Platz 1, 76344 Eggenstein-Leopoldshafen, Germany}
\affil[3]{Department of Aerospace and Mechanical Engineering, University of Southern California, \protect \\3650 McClintock Ave, Los Angeles, CA 90089, USA}
\affil[4]{Department of Chemical Engineering and Materials Science, University of Southern California,\protect \\ 925 Bloom Walk, HED 216, Los Angeles, CA 90089, USA}
\affil[5]{Lawrence Berkeley National Laboratory, Molecular Foundry Division,\protect \\ 1 Cyclotron Rd., Berkeley, 94720 CA, USA}

\begin{document}

\maketitle
\begin{abstract}
\input{00_abstract}
\end{abstract}

\section{Motivation}\label{sec:motivation}
\input{01_motivation}

\section{Existing datasets}\label{sec:existing_datasets}
\input{02_existing_datasets}

\section{Our dataset}\label{sec:our_dataset}
\input{03_our_dataset}

\section{Contributions}\label{sec:contributions}
\input{04_contributions}

\printnomenclature

\clearpage

%\listofchanges
%\clearpage

%\nocite{*}
%\TheBibliography
%\printbibliography[heading=bibintoc]

\bibliography{bib}
\bibliographystyle{rsc} %the RSC's .bst file

\end{document}