### 2024-04-28 paper/02_existing_datasets.tex

% Source of png : https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement

% Zenodo data e.g. is not uniform, easily redable but we aimt ot be
% uniform, easily readable