### 2024-04-28 paper/03_our_dataset.tex

% Sharing raw powder x-ray diffraction data: Challenges and benefits https://www.researchgate.net/publication/329256004_Sharing_powder_diffraction_raw_data_Challenges_and_benefits

% The subsections of this section should be the 
% individual contributions that we get from labs