### 2024-04-29 paper/02_existing_datasets.tex
Included notes on existing datasets

% Source of png : https://www.sciencedirect.com/topics/biochemistry-genetics-and-molecular-biology/rietveld-refinement

% Sources to mention
% Zenodo -> A good amount of raw powder data there but spread into smaller datasets; Inhomogenous, not specially prepared to be worked with by others, therefore difficult to work with especially at scale

% Crystal structures with corresponding simulated diffractograms;
% ->  ICSD (https://icsd.products.fiz-karlsruhe.de/) 
% - 230k structures;
% - Proprietary

% -> PDF (https://www.icdd.com/) -> 
%- 1200k "material data sets" (structures?) 
% + 19 k entries with "Experimental raw data digital patterns"; 
%  Proprietary
%The PDF itself lists the following data sources:
\begin{itemize}
    \item \textbf{ICDD Powders (00) – International Centre for Diffraction Data Newtown Square, PA, USA}
    \item \textbf{ICSD (01) – FIZ Karlsruhe – Leibniz Institute for Information Infrastructure (FIZ), Karlsruhe, Germany – Inorganic Crystal Structure Database}
    \item \textbf{NIST (03) – National Institute of Standards and Technology, Gaithersburg, Maryland, USA}
    \item \textbf{MPDS (04) – Material Phases Data System, Vitznau, Switzerland – Linus Pauling File (LPF)}
    \item \textbf{ICDD Single Crystal Data (05) – International Centre for Diffraction Data Newtown Square, PA, USA}
\end{itemize}

% ->  COD (https://www.crystallography.net/cod/) -
% -  513k structures,
% - Fully open