### 2024-12-01 chapters/06_summary_and_outlook.tex

With the opXRD dataset, a curation of XXX unlabeled and YYY labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns. With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches. The data can be used for the actual method development and for testing. Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.

\paragraph{How to contribute}

Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data. We invite everyone in possession of experimental powder XRD data to submit it to our dataset, to further improve the dataset and aid further developments in this field. Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data. We will keep updating and maintaining the dataset with incoming new submissions.

%We have launched Xqueryer (\url{http://xqueryer.caobin.asia/}), a cutting-edge, free online platform for intelligent structure identification. Users can upload experimental pXRD data to identify crystal structures and extract all relevant information available in the Materials Project database. All usage complies with Xqueryer's terms of service. Uploaded data are securely stored and integrated into the opXRD database.