### 2024-12-19 main.tex

\documentclass[a4paper]{article}

\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[super,sort&compress,comma, square]{natbib}
%\renewcommand{\supercite}[1]{\cite{#1}}

%% Language and font encodings
\usepackage{tikz}
\usepackage{tablefootnote}
\usepackage{gensymb}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{csquotes}
\usepackage{doi}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}

% Useful packages
\usepackage[version=4]{mhchem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue}
\usepackage{authblk}
\usepackage{multirow}
\usepackage{nicematrix,enumitem,booktabs}
\usepackage{nomencl}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{amsfonts}
\usepackage{outlines} %%same
\usepackage{subfig}
\usepackage[version=4]{mhchem}
\usepackage{eurosym}
\usepackage[locale = DE]{siunitx}
\usepackage{bm}
\usepackage[normalem]{ulem} %%temporary, for organization purposes
\usepackage{pythonhighlight}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{adjustbox}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}{\textcolor{green!80!black}{\ding{52}}}
\newcommand{\xmark}{\textcolor{red}{\ding{53}}}
\newcommand{\partialcheck}[1]{\textcolor{orange}{\ding{52}(#1\%)}}
\captionsetup[table]{
  font=footnotesize
}

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\renewcommand\Authfont{\small\raggedright\itshape}  
\renewcommand\Affilfont{\normalfont\small}

\DeclareSIUnit{\myeuro}{\text{\euro}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage[final]{changes}

\interfootnotelinepenalty=10000
\title{opXRD: Open Experimental Powder X-ray Diffraction Database}

\newcommand{\ITI}{1}
\newcommand{\INT}{2}
\newcommand{\HKUSTGZ}{3}
\newcommand{\USC}{4}
\newcommand{\LBLMF}{5}
\newcommand{\LBLALS}{6}
\newcommand{\Hoffmann}{7}
\newcommand{\LBLCS}{8}
\newcommand{\EMPA}{9}
\newcommand{\KITICRT}{10}
\newcommand{\CPT}{11}
\newcommand{\KFUPMMSE}{12}
\newcommand{\KFUPMIRCIMR}{13}
\newcommand{\FZJIEMD}{14}

\author[\ITI,\INT]{Daniel Hollarek}
\author[\ITI,\INT]{Henrik Schopmans}
\author[\ITI,\INT]{Jona Östreicher}
\author[\ITI,\INT]{Jonas Teufel}
\author[\HKUSTGZ]{Bin Cao}
\author[\INT]{Simon Schweidler}
\author[\USC]{Adie Alwen}
\author[\LBLMF]{Mriganka Singh}
\author[\LBLMF,\LBLALS]{Tim Kodalle}
\author[\Hoffmann]{Hanlin Hu}
\author[\LBLCS]{Gregoire Heymans}
\author[\EMPA]{Alexander Wieczorek}
\author[\EMPA]{Siarhei Zhuk}
\author[\CPT]{Arthur Hardiagon}
\author[\KFUPMMSE,\KFUPMIRCIMR]{Maged Abdelsamie}
\author[\HKUSTGZ]{Longhan Zhang}
\author[\HKUSTGZ]{Tong-yi Zhang}
\author[\USC]{Andrea M. Hodge}
\author[\FZJIEMD]{Ruth Schwaiger}
\author[\INT]{Ben Breitung}
\author[\EMPA]{Sebastian Siol}
\author[\KITICRT]{Moritz Wolf}
\author[\CPT]{François-Xavier Coudert}
\author[\LBLMF]{Carolin M. Sutter-Fella}
\author[\ITI,\INT]{Pascal Friederich$^\ast$}

\affil[\ITI]{Institute of Theoretical Informatics, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany.  E-mail: pascal.friederich@kit.edu}
\affil[\INT]{Institute of Nanotechnology, Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany}
\affil[\HKUSTGZ]{Guangzhou Municipal Key Laboratory of Materials Informatics, Hong Kong University of Science and Technology (Guangzhou), Guangzhou 511400, China}
\affil[\USC]{Department of Chemical Engineering and Materials Science, University of Southern California, Los Angeles CA 90089, USA}
\affil[\LBLMF]{Lawrence Berkeley National Laboratory, Molecular Foundry Division, Berkeley 94720 CA, USA}
\affil[\LBLALS]{Lawrence Berkeley National Laboratory, Advanced Light Source, Berkeley 94720 CA, USA}
\affil[\Hoffmann]{Hoffmann Institute of Advanced Materials, Shenzhen Polytechnic, Shenzhen 518055, China}
\affil[\LBLCS]{Lawrence Berkeley National Laboratory, Chemical Sciences Division, Berkeley 94720 CA, USA}
\affil[\EMPA]{Empa–Swiss Federal Laboratories for Materials Science and Technology, 8600 Dübendorf, Switzerland}
\affil[\KITICRT]{Engler-Bunte-Institut \& Institute of Catalysis Research and Technology, Karlsruhe Institute of Technology, Karlsruhe, Germany}
\affil[\CPT]{Chimie ParisTech, PSL University, CNRS, Institut de Recherche de Chimie Paris, 75005 Paris, France}
\affil[\KFUPMMSE]{Material Science and Engineering Department, King Fahd University of Petroleum and Minerals (KFUPM), Dhahran 31261, Saudi Arabia}
\affil[\KFUPMIRCIMR]{Interdisciplinary Research Center for Intelligent Manufacturing and Robotics, KFUPM, Dhahran 31261, Saudi Arabia}
\affil[\FZJIEMD]{Institute of Energy Materials and Devices, Forschungszentrum Juelich GmbH, 52425 Juelich, Germany}

\begin{document}

\maketitle
\textbf{Keywords}: powder x-ray diffraction, machine learning, open-access database, crystal structure determination, phase identification

\begin{abstract}
\input{chapters/00_abstract}
\end{abstract}

\newpage
\section{Introduction}\label{sec:Introduction}
\input{chapters/01_introduction}
\input{chapters/02_SOTA_ML}

\section{Existing datasets}
\input{chapters/03_existing_datasets}

\pagebreak
\section{opXRD database}\label{sec:our_dataset}
\input{chapters/04_our_dataset}

\section{Usage}\label{sec:how_to_use}
\input{chapters/05_how_to_use}

\section{Summary and Outlook}\label{sec:summary_and_outlook}
\input{chapters/06_summary_and_outlook}

\subsection*{Supporting Information}
Supporting Information is available from the Wiley Online Library or the author.

\subsection*{Data availability}
The opXRD database is available on Zenodo at \url{https://zenodo.org/records/14254270}. It is published under the Creative Commons Attribution 4.0 International license.\footnote{Detailed terms of this license can be found at \url{https://creativecommons.org/licenses/by/4.0/deed.en}.} It can be downloaded by any user without any barriers or restrictions. For further details, please refer to Section~$\eqref{sec:how_to_use}$.

\subsection*{Conflicts of interest}
There are no conflicts of interest to declare.

\subsection*{Acknowledgements}
We acknowledge financial support by the German Research Foundation (DFG) through the Research Training Group 2450 “Tailored Scale-Bridging Approaches to Computational Nanoscience”. We acknowledge support by the Federal Ministry of Education and Research (BMBF) under Grant No. 01DM21001B (German-Canadian Materials Acceleration Center). We acknowledge financial support from the Helmholtz Foundation Model Initiative within Project "SOL-AI". Part of this work was funded under the France 2030 framework by Agence Nationale de la Recherche (project ANR-22-PEXD-0009 of PEPR DIADEM). Work at the Molecular Foundry was supported by the Office of Science, Office of Basic Energy Sciences, of the U.S. Department of Energy under Contract No. DE-AC02-05CH11231. Work at the Advanced Light Source (ALS) was done at beamline 12.3.2. The ALS is a DOE Office of Science User Facility under contract no. DE-AC02-05CH11231. The development of the online phase identification platform is supported by the Guangzhou-HKUST(GZ) Joint Funding Program (No. 2023A03J0003). Work by the USC group was supported by the National Science Foundation (NSF) grant numbers DMR-2227178 and OISE-2106597.

%\subsection*{Table of contents}
%\tableofcontents

%A short text and graphic should be provided for the Table of Contents (ToC). The ToC text should describe the main results in 50 to 60 words. It should be written for a general audience and be written in the third person.

%The ToC figure should convey the main message of the article. It does not have to be a figure from the article; it can be a combination of figures or a new, original figure composed to represent the topic. The author must be the copyright holder for this figure and any images used to create it. The size of the image should be either 55 mm × 50 mm (w × h) or 110 mm × 20 mm (w × h).

\printnomenclature

\clearpage

\bibliographystyle{bibstyle/bibstyle.bst}
\bibliography{references}

\end{document}