### 2025-03-01 chapters/06_summary_and_outlook.tex

With the opXRD dataset, a curation of $\numpatterns$ unlabeled and 2179 at least partially labeled experimental powder XRD patterns from a wide range of different materials systems, we provide the largest currently available source of experimental XRD patterns. With this, we address the need for experimental data that arises when developing algorithms and analysis tools for XRD data, both based on machine learning and classical approaches. The data can be used for the actual method development and for testing. Our dataset is a valuable and so far missing resource to drive further developments in the automated analysis of XRD data.

% How to contribute
Rather than a finished project, the opXRD dataset is an ongoing effort to collect experimental powder XRD data. We invite everyone who is working in the area of experimental powder XRD to submit it to the dataset, in order to further improve the utility of the dataset and thus aid further developments in this field. Our submission page (\url{https://xrd.aimat.science/}) and submission helper software will be kept available to collect more data. We will keep updating and maintaining the dataset with new incoming submissions.