### 2025-03-07 chapters/03_existing_datasets.tex

To contextualize opXRD within the current environment of experimental powder diffraction data, the list below provides an overview of the largest crystal structure databases that offer access to experimental powder diffraction data. For an overview of these databases refer to Tab.~$\eqref{tab:exp_databases}$ below.

\begin{table}[!htb]
\centering
\caption{\footnotesize Overview of experimental powder diffraction databases: The column ``O.A.'' indicates whether or not the database is open-access. The availability of the chemical composition, spacegroups, lattice parameters, and atomic coordinates of the underlying samples are indicated by the columns ``Comp.'', ``Spg.'', ``Lattice'' and ``Atom coords.'', respectively.}
\label{tab:exp_databases}
\scalebox{0.82}{
\begin{adjustbox}{center}
\begin{NiceTabular}{@{}llllllll@{}}
\toprule
\textbf{Name} & \textbf{No. patterns} & \textbf{O.A.} & \textbf{Comp.} & \textbf{Spg.} & \textbf{Lattice} & \textbf{Atom coords.} & \textbf{Year est.} \\
\midrule
Linus Pauling file                     & 21,700                    & \xmark          & \cmark        & \cmark        & \cmark        & \cmark          & 2002            \\
Powder Diffraction File \tabularnote[a)]{The PDF lists the Material Platform for Data Science (MPDS) as a database source. Since the MPDS is hosted by the Pauling File project, there is likely significant overlap in the experimental patterns available in the PDF and the Linus Pauling File.} & 20,800 & \xmark & \cmark & \cmark & \cmark & \partialcheck{52} & 1941 \\
RRUFF                                  & 1290                      & \cmark          & \cmark        & \cmark        & \cmark        & \xmark          & 2006            \\
Crystallography Open Database          & 1052                      & \cmark          & \cmark        & \cmark        & \partialcheck{85}        & \partialcheck{85}          & 2003            \\
PowBase                                & 169                       & \cmark          & \cmark        & \xmark        & \xmark        & \xmark          & 1999            \\
\bottomrule
\end{NiceTabular}
\end{adjustbox}}
\end{table}

\textbf{Linus Pauling File}:\cite{PaulingWeb} The Linus Pauling File is a largely commercial crystal structure database published and maintained by the Pauling File project \cite{villars2018}. It is currently distributed as Pearson Crystal data \cite{PearsonWeb} and the Materials Platform for Data Science (MPDS)\cite{MPDSWeb}. The database, first published in 2002, currently contains more than 534,000 crystal structures\cite{MPDSWeb} and 21,700 corresponding experimental powder diffraction patterns\cite{PearsonWeb}. 
This makes the Pauling file, to the best of our knowledge, the largest collection of experimental powder diffraction data available to researchers. As of November 2024, Pearson's crystal data is available to researchers through a purchase of a one-year license starting at a price point of \qty{2200}{\myeuro}\cite{PearsonBuy}. The MPDS is partially open, with the open portion of the MPDS data accessible through a web interface\cite{MPDSWeb}. API access to the full MPDS can be purchased through a one-year license starting at \qty{9500}{\myeuro}\cite{MPDSBuy}. We asked the Pauling File project whether the experimental powder diffraction data is accessible through the MPDS API. The Pauling File project responded that this data is not currently provided through the API, but could be offered in the future at the request of customers.

\textbf{Powder Diffraction File:} \cite{PDFWeb} The Powder Diffraction File (PDF), published and maintained by the International Center for Diffraction Data (ICDD), is a large collection of materials with accompanying powder diffraction data first published in 1941\cite{GatesRector2019}. According to the ICDD the latest release of the PDF, the PDF5+, contains over a million materials with accompanying powder diffraction data. However, since most of these powder diffraction patterns are simulated we asked the ICDD about the number of experimental diffraction patterns in the PDF5+. We were told that 20,800 of the patterns in the PDF5+ stemmed from experiments and that 10,954 of these patterns were accompanied by the atomic coordinates of the underlying structures. Since the PDF5+ lists the MPDS as a database source, there is likely a significant overlap in the experimental patterns found in the PDF5+ and those found in the Pauling file. Currently, the PDF5+ is available to researchers through a purchase of a one-year license starting at a price point of \$6265. However, the ICDD does not allow researchers to train machine learning models on PDF5+ data, regardless of whether the resulting models are published \cite{PDFLicenseWeb}.

\textbf{RRUFF}: \cite{RRUFFWeb} The RRUFF Mineral Database, first published in 2006, provides detailed information on minerals, including their chemical compositions, crystallography, and spectroscopic data \cite{lafuente2015}. Managed by the University of Arizona, it was created to serve as a public repository for mineral identification and research. It contains \num{1290} powder diffraction patterns stemming from experiments each labeled with the lattice parameters and composition of the underlying structures. The RRUFF data is openly accessible on its official website \cite{RRUFFWeb}.

\textbf{Crystallography Open Database:} \cite{CODWeb} The Crystallography Open Database (COD) is an open-access collection of crystal structures founded in 2003\cite{Graulis2009cod}. It currently provides over 500,000 crystal structures. Of these files, 1052 contains the experimental powder diffraction data that was used to determine the underlying crystal structures of the investigated samples. Hence, the experimental powder diffraction data contained in the COD is mostly labeled with the full crystal structure information. The data is openly accessible in the form of .cif files on the official COD website\cite{CODWeb}.

\textbf{PowBase:} \cite{PowBaseWeb} PowBase is a database of 169 mostly unlabeled experimental powder diffraction patterns collected and maintained by crystallography researcher Armel Le Bail starting in 1999. PowBase is an initiative suggested in the Structure Determination by Powder Diffractometry (SDPD) mailing list which was co-maintained by Le Bail. The COD is another community initiative that grew out of this mailing list. As of March 2025, all 169 patterns are still freely available for download on the official website \cite{PowBaseWeb}.

There is also publicly available powder diffraction data uploaded to datasets on Zenodo. However, this data is split into disparate entries that typically only contain the work of a single research project. Additionally, extracting powder diffraction data at scale is hindered by the fact that the data is often given in plain text files in non-standardized formats, which are difficult to parse automatically. We are currently planning a systematic large-scale extraction of powder diffraction data from databases like Zenodo with the help of a large language model. This data will be included in a future release of the opXRD database.

While not strictly speaking a powder diffraction database, the High-Throughput Experimental Materials Database (HTEM) by the National Renewable Energy Laboratory (NREL) is a valuable source of X-ray diffraction data \cite{zakutayev2018}. Currently, the HTEM database contains 65,779 thin-film samples with corresponding X-ray diffraction data\cite{HTEMWeb}. Each database entry includes the elemental composition of the underlying sample but does not provide any information on its structure. HTEM data is open-access and can be downloaded through an API provided by NREL.

Aside from the databases mentioned above, we have also investigated several other crystal structure resources in search of experimental powder diffraction data. Crystal structure resources that were investigated but not found to contain any appreciable amount of publicly available experimental powder diffraction data include the Inorganic Crystal Structure Database \cite{ICSDWeb}, the Cambridge Structural Database \cite{CambridgeWeb}, the Materials Project database \cite{MatProjWeb}, the Crystallographic and Crystallochemical Database \cite{CrystallochemicalWeb}, the Bilbao Incommensurate Crystal Structure Database \cite{BilbaoWeb}, the Mineralogy Database \cite{MineralogyWeb}, the IUCr Raw data letters \cite{IUCrWeb}, the U.S. Naval Research Laboratory Crystal Lattice-Structures \cite{NRLWeb}, the Athena Mineral database \cite{AthenaWeb} and the Protein data bank\cite{PDBWeb}. The lack of experimental powder diffraction data in these databases is to be expected as most structure solutions are achieved through single-crystal diffraction.