### 2025-03-07 chapters/05_how_to_use.tex

The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14254270}) and can be downloaded by any user without any barriers or restrictions.

Next to the availability of the opXRD dataset on Zenodo, we also provide a Python library ``opxrd'' to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library. The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}. The opxrd library includes options for data-loading, standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook (\url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb})
that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.