### 2025-04-28 SI.tex
Added .json description

\documentclass[a4paper]{article}

\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[super,sort&compress,comma, square]{natbib}
%\renewcommand{\supercite}[1]{\cite{#1}}

%% Language and font encodings
\usepackage{tikz}
\usepackage{tablefootnote}
\usepackage{gensymb}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{csquotes}
\usepackage{doi}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{titling}

% Useful packages
\usepackage[version=4]{mhchem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue}
\usepackage{authblk}
\usepackage{multirow}
\usepackage{nicematrix,enumitem,booktabs}
\usepackage{nomencl}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{amsfonts}
\usepackage{outlines} %%same
\usepackage{subfig}
\usepackage[version=4]{mhchem}
\usepackage{eurosym}
\usepackage[locale = US]{siunitx}
\usepackage{bm}
\usepackage[normalem]{ulem} %%temporary, for organization purposes
\usepackage{pythonhighlight}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{parskip}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}[1][]{\textcolor{green!80!black}{#1\quad\ding{52}}}
\newcommand{\xmark}[1][]{\textcolor{red}{#1\quad\ding{53}}}
\newcommand{\partialcheck}[1]{\textcolor{orange}{\quad\ding{52}(#1\%)}}
\captionsetup[table]{font=footnotesize}

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\renewcommand\Authfont{\small\raggedright\itshape}  
\renewcommand\Affilfont{\normalfont\small}
\newcommand{\numpatterns}{92,552 }
\newcommand{\numlabeled}{2179 }

\DeclareSIUnit{\myeuro}{\text{\euro}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage{changes}

\interfootnotelinepenalty=10000
\title{Supporting information for opXRD: Open Experimental Powder X-ray Diffraction Database}

\author[ ]{Daniel Hollarek}
\author[ ]{Henrik Schopmans}
\author[ ]{Jona Östreicher}
\author[ ]{Jonas Teufel}
\author[ ]{Bin Cao}
\author[ ]{Adie Alwen}
\author[ ]{Simon Schweidler}
\author[ ]{Mriganka Singh}
\author[ ]{Tim Kodalle}
\author[ ]{Hanlin Hu}
\author[ ]{Gregoire Heymans}
\author[ ]{Maged Abdelsamie}
\author[ ]{Alexander Wieczorek}
\author[ ]{Siarhei Zhuk}
\author[ ]{Arthur Hardiagon}
\author[ ]{Ruth Schwaiger}
\author[ ]{François-Xavier Coudert}
\author[ ]{Moritz Wolf}
\author[ ]{Sebastian Siol}
\author[ ]{Carolin M. Sutter-Fella}
\author[ ]{Ben Breitung}
\author[ ]{Andrea M. Hodge}
\author[ ]{Tong-yi Zhang}
\author[*]{Pascal Friederich}

\affil[*]{Corresponding author: pascal.friederich@kit.edu}

\begin{document}

\maketitle
\renewcommand{\thesection}{S\arabic{section}}

\section{Description of opXRD files on Zenodo}
\replaced{The opXRD database is packaged into a single zip file ``opxrd.zip'' available for download on Zenodo.}{The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.} Within the .zip archive patterns are saved as .json files grouped in folders indicating the contributing institution. If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project. These research project folders are labeled alphabetically in the order they are introduced in Section~$3$. \added{Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series. }

Each .json file contains a pattern recorded from an X-ray diffraction experiment\added{, as well as any available structural information of the underlying sample}. \added{At its topmost layer the .json file contains 4 mandatory fields: ``two\_theta\_values'', ``intensities'', ``powder\_experiment'' and ``metadata''. The first two together define the content of the diffraction pattern. The powder experiment field contains all available details about the experiment, including the wavelength of the employed X-ray source and everything known about the sample underlying the experiment. Insofar as these properties are available, stored sample properties include the chemical composition, lattice parameters, space group, atomic coordinates, and phase fraction of each phase. Finally, the metadata field contains information about the origin of the data including contributing institution, the name of the contributor as well as the original file format and measurement date if available.}

\added{For further details on the structure of the JSON file, we refer to the custom Python class ``XrdPattern'' in the \textsc{xrdpattern} library. Since the .json file serializes the ``XrdPattern'' class, the fields of the .json file are in direct correspondence with the attributes of the python class.}\deleted{If available, the composition and structure of the investigated sample and experiment conditions are also included in this file. Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.}

\section{opXRD Python library \deleted{usage}}

\added{Next to the dataset on Zenodo we also offer a dedicated Python library interface with the dataset directly through Python.} The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}. If the database is locally available under \pyth{root_dirpath} this command loads the library from this location. If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}. \added{The opXRD library also contains the code used to generate the tables and figures of this work: \url{https://github.com/aimat-lab/opXRD/blob/main/opxrd/plots.ipynb}.}

\pagebreak

\section{Combined pattern plots}
Figure~$\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/combined.png}
    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset. The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
    \label{fig:combined}
\end{figure}

\end{document}