### 2025-04-28 chapters/05_how_to_use.tex

The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14254270}) and can be downloaded by any user without any barriers or restrictions.

\added{The opXRD library is designed for easy integration with common machine learning frameworks such as \emph{PyTorch}. This makes it an ideal resource for researchers developing and benchmarking sim-to-real transfer strategies in pXRD data analysis.}
\replaced{We}{Next to the availability of the opXRD dataset on Zenodo, we also} provide a Python library ``opxrd'' to easily download and interface with the dataset. The instructions for how to install this library can be found in the repository associated with the library. The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}. The opxrd library includes options for data-loading, standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook (\url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb})
that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.