### 2025-04-29 SI.tex
Finalized changes @ 14:41 29.04.25

\documentclass[a4paper]{article}

\pdfoutput=1
\usepackage[utf8]{inputenc}
\usepackage[super,sort&compress,comma, square]{natbib}
%\renewcommand{\supercite}[1]{\cite{#1}}

%% Language and font encodings
\usepackage{tikz}
\usepackage{tablefootnote}
\usepackage{gensymb}
\usepackage[english]{babel}
\usepackage[T1]{fontenc}
\usepackage{csquotes}
\usepackage{doi}
%% Sets page size and margins
\usepackage[a4paper,top=3cm,bottom=2cm,left=3cm,right=3cm,marginparwidth=1.75cm]{geometry}
\usepackage{titling}

% Useful packages
\usepackage[version=4]{mhchem}
\usepackage{amsmath}
\usepackage{graphicx}
\usepackage{hyperref}
\hypersetup{colorlinks=true, allcolors=blue}
\usepackage{authblk}
\usepackage{multirow}
\usepackage{nicematrix,enumitem,booktabs}
\usepackage{nomencl}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{textcomp}
\usepackage{amsfonts}
\usepackage{outlines} %%same
\usepackage{subfig}
\usepackage[version=4]{mhchem}
\usepackage{eurosym}
\usepackage[locale = US]{siunitx}
\usepackage{bm}
\usepackage[normalem]{ulem} %%temporary, for organization purposes
\usepackage{pythonhighlight}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{adjustbox}
\usepackage{parskip}

\usepackage{pifont}% http://ctan.org/pkg/pifont
\newcommand{\cmark}[1][]{\textcolor{green!80!black}{#1\quad\ding{52}}}
\newcommand{\xmark}[1][]{\textcolor{red}{#1\quad\ding{53}}}
\newcommand{\partialcheck}[1]{\textcolor{orange}{\quad\ding{52}(#1\%)}}
\captionsetup[table]{font=footnotesize}

\date{}
\newcommand{\red}[1]{\textcolor{red}{#1}}
\renewcommand\Authfont{\small\raggedright\itshape}  
\renewcommand\Affilfont{\normalfont\small}
\newcommand{\numpatterns}{92,552 }
\newcommand{\numlabeled}{2179 }

\DeclareSIUnit{\myeuro}{\text{\euro}}
\DeclareSIUnit\angstrom{\text {Å}}
\usepackage{changes}

\interfootnotelinepenalty=10000
\title{Supporting information for opXRD: Open Experimental Powder X-ray Diffraction Database}

\author[ ]{Daniel Hollarek}
\author[ ]{Henrik Schopmans}
\author[ ]{Jona Östreicher}
\author[ ]{Jonas Teufel}
\author[ ]{Bin Cao}
\author[ ]{Adie Alwen}
\author[ ]{Simon Schweidler}
\author[ ]{Mriganka Singh}
\author[ ]{Tim Kodalle}
\author[ ]{Hanlin Hu}
\author[ ]{Gregoire Heymans}
\author[ ]{Maged Abdelsamie}
\author[ ]{Alexander Wieczorek}
\author[ ]{Siarhei Zhuk}
\author[ ]{Arthur Hardiagon}
\author[ ]{Ruth Schwaiger}
\author[ ]{François-Xavier Coudert}
\author[ ]{Moritz Wolf}
\author[ ]{Sebastian Siol}
\author[ ]{Carolin M. Sutter-Fella}
\author[ ]{Ben Breitung}
\author[ ]{Andrea M. Hodge}
\author[ ]{Tong-yi Zhang}
\author[*]{Pascal Friederich}

\affil[*]{Corresponding author: pascal.friederich@kit.edu}

\begin{document}

\maketitle
\renewcommand{\thesection}{S\arabic{section}}

\section{Description of opXRD files on Zenodo}
\replaced{The opXRD database is packaged as a single zip file ``opxrd.zip'' available for download on Zenodo.}{The database comes in two zip archives, ``opxrd.zip'' and ``opxrd\_in\_situ.zip''. The latter contains the in-situ data with highly correlated patterns recorded through time series measurements.} Within the .zip archive patterns are saved as .json files grouped in folders indicating the contributing institution. If an institution contributed data from several projects, the contributed data is further divided into folders indicating the research project. These research project folders are labeled alphabetically in the order they are introduced in Section~$3$. \added{Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series. }

Each .json file contains a pattern recorded from an X-ray diffraction experiment\added{, as well as any available structural information of the underlying sample}. \added{At its topmost layer, the .json file contains 4 mandatory fields: ``two\_theta\_values'', ``intensities'', ``powder\_experiment'', and ``metadata''. The first two define the content of the diffraction pattern. The ``powder\_experiment'' field contains all available details about the experiment, including the wavelength of the employed X-ray source and everything known about the sample underlying the experiment. Insofar as these properties are available, stored sample properties include the chemical composition, lattice parameters, space group, atomic coordinates, and the phase fraction of each phase of the material. Finally, the metadata field contains information about the origin of the data, including the contributing institution, the name of the contributor as well as the original file format and measurement date if available. The metadata field will in the future also display the version of the opXRD database in which the pattern was added.}

\added{For further details on the structure of the JSON file, we refer to the custom Python class ``XrdPattern'' in the \textsc{xrdpattern} library. Since the .json file serializes the ``XrdPattern'' class, the fields of the .json file are in direct correspondence with the attributes of the python class.}\deleted{If available, the composition and structure of the investigated sample and experiment conditions are also included in this file. Patterns belonging to time series measurements are labeled with filenames that indicate the measurement series they belong to and their order in that series.}

\section{Figure source code\deleted{opXRD Python library usage}}

\deleted{Next to the dataset on Zenodo we also offer a dedicated Python library interface with the dataset directly through Python. The opXRD Python library allows the dataset to be accessed through one simple command: \pyth{OpXRD.load(root_dirpath)}. If the database is locally available under \pyth{root_dirpath} this command loads the library from this location. If the database is not available locally at this location, the database is automatically downloaded from Zenodo to \pyth{root_dirpath}.}
\added{The opXRD library also contains the source code that was used to generate the tables and figures seen in Section$~3$ and the appendix. All figures, including the associated code to generate them, are located in the following notebook: \url{https://github.com/aimat-lab/opXRD/blob/9a6050ea8b62c9d67f72a126a33bf18f70cebefb/opxrd/plots.ipynb}.}

\pagebreak

\section{Combined pattern plots}
Figure~$\eqref{fig:combined}$ shows 50 randomly selected samples of the X-ray diffraction patterns found in each of the research projects contributed to the opXRD database.

\begin{figure}[!htb]
    \centering
    \includegraphics[width=0.95\linewidth]{figures/combined.png}
    \caption{50 randomly chosen X-ray diffraction patterns from each contributed dataset. The figure shows data from the following datasets: a) EMPA, b) LBNL-A, c) LBNL-B, d) LBNL-C, e) USC, f) INT, g) HKUST-A, h) HKUST-B, i) CNRS, j) IKFT.}
    \label{fig:combined}
\end{figure}

\end{document}