### 2025-04-29 chapters/05_how_to_use.tex

The opXRD database is hosted on Zenodo (\url{https://zenodo.org/records/14254270}) and can be downloaded by any user without any barriers or restrictions.

\replaced{We}{Next to the availability of the opXRD dataset on Zenodo, we} also provide a Python library ``opxrd'' to easily download and interface with the dataset.
\added{The opXRD library is designed for easy integration with common machine learning frameworks such as \emph{PyTorch}. This makes it an ideal resource for researchers developing and benchmarking sim-to-real transfer strategies in pXRD data analysis.}
The instructions for how to install this library can be found in the repository associated with the library. The repository to this library is located at \url{https://github.com/aimat-lab/opxrd}. The opxrd library includes options for data-loading, standardization, plotting, and the conversion to \emph{PyTorch} tensors. We provide a Jupyter Notebook (\url{https://colab.research.google.com/github/aimat-lab/opXRD/blob/main/opxrd/usage.ipynb})
that showcases these functionalities in more detail. This notebook also illustrates how to interface with the opXRD database through Python.